Diffusive Homeostasis in a Self-Organizing Recurrent Neural Network \\ \begin{large} Spatially dependent Interaction as a Determinant of Neural Activity and Plasticity \end{large}

### 2016-12-19 draft.tex
.

\section{Methods} \label{methods}
\subsection{Network Simulation}

@@

\subsection{Network Topology}
So far, we have only analyzed network activity, whereas in this section we present results that relate to features of synaptic topology. Since the SORN we used for all simulations only included plastic mechanisms affecting recurrent excitatory connections, these were the main subject of our investigations. Moreover, in contrast to the previous section, we did not explicitly aim for the emergence of an explicitly \emph{new} feature of synaptic topology, but rather sought to recover properties that had been found in earlier versions of the SORN.

>>>>

\section{Methods} \label{methods}
\subsection{Network Simulation} \label{network simulation}

@@

\subsection{Network Topology}
So far, we have only analyzed network activity, whereas in this section we present results that relate to features of synaptic topology. Since the SORN we used for all simulations only included plastic mechanisms affecting recurrent excitatory connections, these were the main subject of our investigations. Moreover, in contrast to the previous section, we did not aim for the emergence of an explicitly \emph{new} feature of synaptic topology, but rather sought to recover properties that had been found in earlier versions of the SORN.

\subsubsection{Excitatory Connection Fraction}
In the previous version of the SORN, initializing the network with zero recurrent excitatory connections led to a monotonically increasing but saturating excitatory-to-excitatory connection fraction (CF). The growth rate was tuned such that the terminal CF settled at $\mathrm{10\%}$, see \cite[p. 8]{SORN_Paper}. Omitting the distance dependence of connection probability for synaptic growth led to a slightly higher connection fraction. To compare these previous results to the diffusive case, we ran a full simulation under diffusive homeostasis, including the growth phase by setting the target concentration to an appropriate value that was determined in a previous simulation with non-diffusive homeostasis. The resulting time course of CF is depicted in Figure \ref{CF_plot}.
\begin{figure}
%%% plot_CF.py
\includegraphics[width=\textwidth]{../../plots/CF_plot.png}
\caption{Connection fractions of recurrent excitatory synapses. Simulation protocols include all four possible combinations of diffusive/non-diffusive homeostasis and distance-dependent/non-distance-dependent synaptic growth, each variant retained for the entire simulation time.}
\label{CF_plot}
\end{figure}
To our surprise, diffusive homeostasis caused the CF to overshoot its final fraction in the growth phase. Eventually however, it settled at a similar equilibrium of $\mathrm{\approx 10\%}$. To understand this kind of behaviour, we recalled the two mechanisms that directly determine changes within the connection fraction, namely synaptic growth and pruning. The average change of connection fraction is simply proportional to the difference between newly created synapses and those that were removed. Though fluctuating, the synaptic growth rate was kept at a constant level of 920 synapses/second. In consequence, any differences in Figure \ref{CF_plot} must originate in different pruning rates. As explained in section \ref{network simulation}, pruning is performed once per second by collection connections that have fallen under the given weight-threshold. Within this second, a synaptic weight can fall below this value for two reasons: either due to multiplicative normalization or additive STDP. If the weight of a neuron is already very small, multiplicative changes due to normalization have a minor effect compared to weight fluctuations caused by additive STDP, drawing us to the conclusion that differences within the statistics of STDP are the potential cause for the observed overshooting of the CF.
